---
title: "Bootstrap Sampling"
author: "SXX-XXX"
date: "2022-11-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(ggplot2)
```

Body fat percentage of 92 individuals are given below. 
```{r}
data <- c(25.3,29.3,37.7,32.8,24.6,26.5,21.2,28.4,24,28.7,37.4,30.7,36.7,28.2,26.4,37.1,
          31.1,43.1,34.1,26.7,30.9,30,22.1,24.4,22.5,24.9,18.2,27.3,25.9,28.3,18.7,22.4,
          23.6,26.8,17.8,27.4,16.8,26.1,20.8,22.7,20.2,20.3,31.9,22.9,25.3,17.3,41,25.8,
          36.6,27.8,39.4,36.4,46.8,40.5,43,39.5,39.4,24.8,35,25.3,42.5,27.8,35.8,39.2,
          38.3,34.1,39.9,32.5,29.9,32,23,28.5,26.4,33.9,29.4,29.4,21,22.4,25.3,20.2,22.5,
          23.4,29.5,21.9,19.2,28.6,23.4,23.2,18.1,31.4,22.9,25.6)
```

1. Draw a histogram of bodyfat data
```{r}

```


2. Check the normality of observations
 
```{r}

```

3. Obtain the sampling distribution of average body fat percentage
```{r}

```


4. Obtain a estimate for average body fat of an individual using Bootstrap sampling 

```{r}
 
```


5. Draw a histogram of bootstrap estimates and sampling distribution of average body fat of an individual
```{r}


```
6. Estimate the median body fat percentage of an individual

```{r}

```
7. Obtain the 95% confidence interval for  median body fat percentage of an individual
```{r}
```
 