---
title: "Bootstrap Sampling"
author: "SXX-XXX"
date: "2022-11-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(ggplot2)
```

Body fat percentage of 92 individuals are given below. 
```{r}
data <- c(25.3,29.3,37.7,32.8,24.6,26.5,21.2,28.4,24,28.7,37.4,30.7,36.7,28.2,26.4,37.1,
          31.1,43.1,34.1,26.7,30.9,30,22.1,24.4,22.5,24.9,18.2,27.3,25.9,28.3,18.7,22.4,
          23.6,26.8,17.8,27.4,16.8,26.1,20.8,22.7,20.2,20.3,31.9,22.9,25.3,17.3,41,25.8,
          36.6,27.8,39.4,36.4,46.8,40.5,43,39.5,39.4,24.8,35,25.3,42.5,27.8,35.8,39.2,
          38.3,34.1,39.9,32.5,29.9,32,23,28.5,26.4,33.9,29.4,29.4,21,22.4,25.3,20.2,22.5,
          23.4,29.5,21.9,19.2,28.6,23.4,23.2,18.1,31.4,22.9,25.6)
```

1. Draw a histogram of body fat data
```{r}
hist(data)
```
Note: 
Based on the histogram it is clear that data is not symmetrically distributed.
Therefore, it's assumed that it is not a normal distribution.

2. Check the normality of observations

Methods of normality test. (https://www.statology.org/test-for-normality-in-r/)
  1. (Visual Method) Create a histogram.
      If the histogram is roughly “bell-shaped”, then the data is assumed to 
      be normally distributed.

  2. (Visual Method) Create a Q-Q plot.
      If the points in the plot roughly fall along a straight diagonal line, 
      then the data is assumed to be normally distributed.

  3. (Formal Statistical Test) Perform a Shapiro-Wilk Test.
      If the p-value of the test is greater than α = .05, then the data is 
      assumed to be normally distributed.

  4. (Formal Statistical Test) Perform a Kolmogorov-Smirnov Test.
      If the p-value of the test is greater than α = .05, then the data is 
      assumed to be normally distributed.
 
```{r}
# Here Q-Q plot is used for normality test
qqnorm(data)
qqline(data)

```

Note: 
Data is not following this diagonal line therefore, it is not a normal distribution. 

3. Obtain the sampling distribution of average body fat percentage
```{r}
# sample mean 
x_bar <- mean(data)
# Variance
x_S_2 <- var(data)


n <- length(data)



# What is the sampling distribution of the x bar
# Based on the CLT, N(x_bar, x_S_2/n)
print(paste("sample size:" , n), quote = FALSE )
print(paste("sample mean:",x_bar), quote = FALSE )
print(paste("sample variance:",x_S_2 / n ), quote = FALSE )



```


4. Obtain a estimate for average body fat of an individual using Bootstrap sampling 

```{r}
set.seed(12345) 
#Bootstrap Samples 
N <- 10000
#vector to hold Bootstrap means 
bt_mean <- rep(NA, N)

#Loop to create N number of samples 
for (i in 1:N)
{
  #create sample 
  bt_sample <- sample(data, n, replace = TRUE)
  bt_mean[i] <- mean(bt_sample)  
}

bootstarp_mean <- mean(bt_mean)
bootstrap_S_2 <- var(bt_mean)

bootstarp_mean
bootstrap_S_2
```


5. Draw a histogram of bootstrap estimates and sampling distribution of average body fat of an individual
```{r}
ggplot(data=data.frame(bt_mean = bt_mean), aes(x=bt_mean))+
  geom_histogram(aes(y = ..density..),
                      bandwidth = 0.5,
                      colour = "black",
                      fill = "white") +
  stat_function(fun = dnorm, args = list(mean = x_bar, sd = sqrt(x_S_2/n)), colour = "blue")+
  stat_function(fun = dnorm, args = list(mean = bootstarp_mean, sd = sqrt(bootstrap_S_2)), colour = "red")+
  theme_classic()

```
5.2 Estimate 95% confidence interval for mean 
```{r}
quantile(bt_mean, c(0.025, 0.975))
```


6. Estimate the median body fat percentage of an individual

```{r}

```
7. Obtain the 95% confidence interval for  median body fat percentage of an individual
```{r}
```
 