---
title: "Factor_Analysis"
author: "Sachith M Gunawardane"
date: "2023-06-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data 
#Data file is available under the same Git folder 
```{r}
data <- read.table("D:/PGIS_Data_Science_AI/DS5110_Statistical_simulation/projects/Factor_Analysis/track_rec.csv", header = T, sep =",")
head(data)

```

```{r}
dim(data)
```
```{r}
dataset1 <- data[,-1]
head(dataset1)
```
#Check for Correlation 
```{r}
cor(dataset1)
```
As you can see data is highly /significantly correlated

#Bartlett.test 
This is to verify is there any possibility to do factor analysis 
```{r}
bartlett.test(dataset1)
```
p-value is small ~ 0; which mean significant  
hence factor analysis is possible 

#Check all variables are good for factor analysis or not 
```{r}
library(psych)
KMO(cor(dataset1))

```
1st - Overall MSA (Measure of Sampling Adequacy) =  0.89
If MSA is < 0.5, it indicate that overall factor analysis is not possible 

2nd - Individual MSA also above 0.5 hence all variables are good for Factor analysis

#FACTOR analysis
```{r}
factor1.out <-  factanal(dataset1, factors = 1 )
factor1.out
```
It is important to look at P-Value 
H0: is One Factor is sufficient 
P-Value < 0.001 hence we reject H0 and conclude that 1 factor is not sufficient for this data set 

```{r}
factor2.out <-  factanal(dataset1, factors = 2 )
factor2.out
```
P-value > 0.001 therefore we fail to reject H0
Hence 2 factor solution is sufficient for this 

factor1 : X1500m & X5000m & X10000m and Marathon. Indicate all for long distance runs 
factor2: 100m, 200m, 400m and 800m which are short distance 

```{r}
factor3.out <-  factanal(dataset1, factors = 2, rotation = "none" )
factor3.out
```
when we set rotation to varimax it try to optimize/maximize the variance 
this is useful when variables are hard/difficult to interpret 

#How to identify # of factors required 
```{r}
eigen(cor(dataset1))
```
# of fators required should eigen values > 1

```{r}
library(factoextra)
pca.out <-  PCA(dataset1, ncp = 8)
fviz_eig(pca.out)
```


